---
title: "CyChecks"
author: "Many"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CyChecks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(CyChecks)
library(dplyr)
library(ggplot2)
```
## Abstract

Salary data for Iowa State University (ISU) employees is publicly available through the state government. CyChecks, an R package, was developed with the intent of examining and interpreting the university's salary dataset. This package will allow users to compare pay between genders, positions, and departments, while displaying the results in a dynamic interface (shiny app). While this data is publicly available, it was also understood that salaries are sensitive, personal information and therefore an individual's name that could easily be associated with their salary was anonymized. We hope Cychecks will provide useful fodder for discussing pay patterns within a publicly funded university.


## Introduction
The goal of CyChecks is to allow users easy access to the publicly available data concerning Iowa State University (the Cyclones) employee salaries.
#### Some of the services this access could offer include:

1.	Aiding job seekers in negotiating starting salaries.
2.	Shedding light on possible pay inequities with respect to gender.
3.	Identifying the highest paid positions at the university.


## Literature Review

Pay patterns and/or pay disparities have been an on-going topic of discussion and focus for research since [insert date]. 
The [insert law] aimed to dismantle these disparities but unfortunately the female gender is still under xx.

Studies have shown gender pay gaps are often a result of gender-associated differences in the  aggressiveness of starting salary negotiations. Transparency about collegue pay could help address this issue.  

## Package Contents

The state of Iowa offers a large amount of public data at this [site](https://data.iowa.gov/). You can access the data by signing up for an API token [here](https://dev.socrata.com/foundry/data.iowa.gov/s3p7-wy6w). Iowa state employee salaries are available at this [site](https://data.iowa.gov/State-Finances/State-of-Iowa-Salary-Book/s3p7-wy6w). Using an API token, CyChecks provides a function to easily get data for any given year.
The data from this site does not include the employee's home department. Unfortunately, after much exploring, the data linking names to departments is not easily accessible. Iowa State Univerisity's Human Resources Department kindly provided a list of employees with their home departments and associated colleges valid as of January 1st 2019. Since acquiring this information is not reproducible to the average user, included is a full dataset (with names anonymized) of all salary info (from years 2007 to 2018) cross-referenced by department in the CyChecks package. It was discovered that employees that had left before the year 2019 will not listed under their respective department in this dataset. This isn't ideal, but a reality of trying to display this sort of data.


## Package Description & Package Functions
 This package (CyChecks) is an R package that helps a user conveniently "check" salary information for Iowa State University employees and faculty. The package currently includes:

1. A function to:
   (i) download data directly from the iowa.gov website
   (ii) anonymize names of downloaded data
   (iii) simplify professor position titles
   (iv) quickly identify departments with possible gender pay inequities
   (v) Launch a Shiny App to help you explore the dataset provided in our package.
2. Data from:
    (i) Employee/salary/department dataset for 2007-2018 
    (ii) Employee/salary/department/college dataset assumed to be valid for 2018
3. A shiny app for visualizing data

## Function 1: Web Scrapping (sal_df)
-Get data from the web
```{r}
#insert example of what function executes

```


## Function 2: Anonymize names (anonymize)
-convert each individual's personal name to an alphanumeric id to mask the individual's real identity
```{r}
#insert example of what function executes

```

## Function 3:Get professor info (get_profs)
-find the pay gap between different positions
```{r}
#insert example of what function executes

```

## Function 4: Run basic statistics (stats_mf)
This function uses one fiscal year of data to identify departments with possible gender pay dispartities. Within a department, the function identifies positions that have more than 1 female and male in that position. Within a department, the function then fits a simple linear model of the following form:
total_salary ~ position + gender
The function then sorts the departments by the p-value associated with the gender term of the linear model, with the lowest p-values appearing first. The function then assigns a verdict, with a p-value less than 0.20 illiciting a 'boo' verdict, and a p-value higher than 0.20 earning an 'ok'. This allows users to quickly filter to find departments with possible pay inequities associated with genders. 

```{r stats_mf ex, eval = F}
data(sals18)
sals18 %>%
  stat_mf() %>%
  filter(verdict == 'boo')
```


### Data

The data *sals18* is included with our package, and it blah blah.  

```{r data ex1}
data("sals18")
names(sals18)
```

## Figures produced with package contents

The following figures are used in the Shiny App to help the user visualize pay patterns. 

This figure demonstrates the 
```{r fig1 ex, echo = F, fig.cap="Professor 2018 salaries by gender and position in the Agronomy Department"}
data("sals_dept_profs") 

sals_dept_profs %>%
   filter(gender %in% c("M", "F"),
          department == "AGRONOMY",
          fiscal_year == 2018) %>%
   ggplot(aes(x = gender, 
              y = total_salary_paid/1000,
              group = position_simplified)) +
        geom_jitter(size = 2, width = 0.1, alpha = 0.5, 
                    aes(color = position_simplified)) +
        stat_summary(fun.y = mean, geom = "line", color = "gray") +
        stat_summary(fun.y = mean, geom = "point", size = 5,  pch = 24,
                     color = "black",
                     aes(fill = position_simplified)) +
        labs(x = NULL, y = "Total Salary Paid\nThousands of $", 
             fill = "Professor Position") +
        theme_bw() +
  guides(color = F) +
        #scale_color_manual(values = c("darkblue", "goldenrod")) +
        theme(#legend.position = c(0.01, 0.99),
              #legend.justification = c(0, 1),
              legend.background = element_rect(linetype = "solid",
                                               color = "black"))
 
```

From this figure you can see that only 3 of the 6 professor positions have both male and females respresented. The gray lines connect mean salaries of each position - you can quickly see there is a negative slope to the lines, indicating females in the same positions on average earn less than their male counterparts in the same positions. You can also see there are less dots in the female category, indicating there are more males in the professor positions compared to females. This type of graph is accessible in our Shiny App **PROF** tab. 

An additional figure that looks at male versus female make-ups of departments is also included in the Shiny App. 

```{r fig2 ex, echo = F, fig.cap="Number and gender of employees in professor positions in the Agronomy Dept. based on departmental affiliations in 2018"}
  sals_dept_profs %>%
   filter(gender %in% c("M", "F"),
          department == "AGRONOMY",
          !is.na(total_salary_paid)) %>%
  group_by(fiscal_year, gender) %>%
        summarise(n = n()) %>%
  ggplot(aes(x = fiscal_year, y = n, fill = gender)) +
      geom_col() +
      theme_bw() +
      scale_fill_manual(values = c("darkblue", "goldenrod")) +
      labs(x = NULL, y = "Number of Employees", fill = "Gender") +
      theme(legend.position = c(0.01,0.99),
            legend.justification = c(0,1),
            legend.background = element_rect(linetype = "solid", color = "black"))

```

From this figure you can see the department has not hired a woman to a professor position since 2014. This is consistent with the previous figure, which showed there were no women in the 'Assistant Professor' position. 



## Conclusion

## Limitations & Future Work

We were only granted access to departmental affiliations for people employed by Iowa State University as of Jan 1 2019. If an employee left the university before that date, they will not be included in our dataset. This makes interpreting post-doctoral information particularly problematic. Despite these limitations, we believe this work opens the door to conversations about pay inequalities at ISU, and can serve as a resouce for individuals looking to negotiate starting salaries or request raises. Future work will include providing a way to quickly identify positions within a department that lack gender diversity, incorporating individuals who do not identify with one of the binary gender categories, and including a way to access departmental affiliations from the web. 

## References/Bibliography

## Packaage Website/Vignette
Follow the [link](https://vanichols.github.io/CyChecks/)
